---
title: "EM-DAT Analysis"
subtitle: "How many disasters have there been?"
author: "Bryan Whiting"
date: "8/17/2021"
output: 
  bookdown::html_document2:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
    fig_caption: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
easypackages::libraries("readxl", "dplyr", "ggplot2", "gt")
theme_set(ggthemes::theme_fivethirtyeight())
```

```{r read_df}
# Load data
# Source: https://public.emdat.be/data
# All Natural, 2001-2021, all countries
path <- "/home/rstudio/data/emdat/emdat_public_2021_08_17_query_uid-84GUwE.xlsx"
# data starts on 6th row
df_raw <- read_excel(path, skip = 6)
df <- df_raw %>%
  janitor::clean_names() %>%
  mutate(year = as.numeric(year))
```

I want to understand if disasters are increase (as they seem to be). The EM-DAT database collects major natural disasters recorded across the world. Not sure if this is the 
best dataset, or if it's the most complete. Also, there's selection bias in what
gets recorded. Also, I'm not an expert on this dataset, so my conclusions could 
be wrong. Poke around for yourself.

# Disasters by year
```{r}
df_tmp <- df %>%
  count(year)

df_tmp %>%
  ggplot(aes(x = year, y = n)) +
  geom_point() +
  geom_smooth(color = "black") +
  labs(
    title = "Count of global natural disasters over time",
    subtitle = "Seem to be trending downward?",
    caption = "Source: EM-DAT database",
    y = "",
    x = ""
  )
```

# Disasters by subgroup
```{r}
df %>%
  count(year, disaster_subgroup) %>%
  ggplot(aes(x = year, y = n, fill = disaster_subgroup)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Disasters by Group",
    fill = "Sub Group",
    caption = "Source: EM-DAT database"
  )

a <- 10
```
